---
title: "Citing R and R Packages"
date: today
format:
  html:
    toc: true
    self-contained: true
---

<!--
  Template: cite-r-packages.qmd
  Purpose:  Generate BibTeX entries and in-text citations for R and the
            R packages used in your analysis. Three approaches are shown,
            from simplest (built-in) to most automated (grateful).

  Usage:
    1. Copy this file into your project directory.
    2. Edit the `pkgs` vector below to list the packages you actually used.
    3. Render with:  quarto render cite-r-packages.qmd
    4. Merge the generated packages.bib entries into your project's
       references.bib (or use packages.bib directly).
-->

# Why Cite R Packages?

Software citation is important for three reasons (LaZerte 2021):
\

1. **Developer credit** — many package authors are scientists who need citations
   for career recognition.
2. **Reproducibility** — recording package names and versions lets others
   replicate your analysis environment.
3. **Gratitude** — citing open-source tools acknowledges unpaid community work.

The [FORCE11 Software Citation Principles](https://doi.org/10.7717/peerj-cs.86)
recommend citing all software important to the research outcome.

# Setup

List the packages you used in your analysis. Edit this vector to match
your actual workflow.

```{r}
#| label: setup
#| message: false

# ── Packages you want to cite ──────────────────────────────────────────
# Edit this list to match your project. Include every package that was
# material to your results (statistical methods, domain-specific tools,
# data retrieval, visualization). General-purpose packages (e.g., dplyr,
# ggplot2) are worth citing too — see guidance below.
pkgs <- c(
  "tidyverse",
  "metafor",
  "ggplot2",
  "dplyr"
)

# Output file for generated BibTeX entries
bib_file <- "packages.bib"
```

# Approach 1: Built-in `knitr::write_bib()` (recommended)
\
This approach requires **no additional packages** beyond `knitr`, which is
already installed if you use Quarto or R Markdown.

```{r}
#| label: knitr-write-bib

# Generate BibTeX entries for R itself plus your packages
knitr::write_bib(c("base", pkgs), file = bib_file)
cat("Wrote", length(c("base", pkgs)), "entries to", bib_file, "\n")
```

The generated `packages.bib` now contains `@Manual{}` or `@Article{}` entries
for each package plus R itself. You can merge these into your project's
`references.bib` or point your Quarto YAML at both files:
\
```yaml
bibliography:
  - references.bib
  - packages.bib
```

## Individual citations

You can also inspect the citation for a single package interactively:

```{r}
#| label: single-citation

# View the citation object for one package
citation("metafor")
```

## BibTeX for a single package

```{r}
#| label: single-bibtex

# Print the BibTeX entry for one package
toBibtex(citation("metafor"))
```

# Approach 2: `grateful` Package (most automated)
\
The [`grateful`](https://github.com/Pakillo/grateful) package scans the
packages loaded in your session and generates a formatted citation paragraph
plus a `.bib` file automatically. Install it once:

```r
install.packages("grateful")
```

```{r}
#| label: grateful-cite
#| eval: false

# Load all the packages you used in your analysis first, then:
library(grateful)

cite_packages(
  output    = "paragraph",
  out.dir   = ".",
  bib.file  = "grateful-refs",
  include.RStudio = FALSE
)
# Creates grateful-refs.bib and prints a formatted paragraph you can
# paste into your Methods section.
```

The paragraph output reads something like:

> "All analyses were performed using R Statistical Software (v4.4.1;
>  R Core Team, 2025). Packages used: dplyr (v1.1.4; Wickham et al., 2023),
>  ggplot2 (v3.5.1; Wickham, 2016), metafor (v4.6-0; Viechtbauer, 2010),
>  tidyverse (v2.0.0; Wickham et al., 2019)."

# Approach 3: Session Information (environment capture)
\
For a complete record of every loaded package and its version, append
`sessionInfo()` or `devtools::session_info()` at the end of your analysis
script. This is essential for reproducibility even if you also cite
individual packages.

```{r}
#| label: session-info

sessionInfo()
```

# Example Methods Paragraph

Adapt the following for your manuscript's Statistical Analysis or Software
subsection:

> All analyses were performed using R Statistical Software
> (v`r paste0(R.version$major, ".", R.version$minor)`;
> R Core Team, `r R.version$year`). Meta-analysis was conducted with the
> metafor package (v`r packageVersion("metafor")`; Viechtbauer, 2010).
> Data manipulation used dplyr (v`r packageVersion("dplyr")`; Wickham
> et al., 2023) and visualization used ggplot2
> (v`r packageVersion("ggplot2")`; Wickham, 2016).

# Merging into Your Project Bibliography

If your project already has a `references.bib`, you can append the
generated package entries:

```{r}
#| label: merge-bib
#| eval: false

# Read the generated package bibliography
pkg_bib <- readLines("packages.bib")

# Append to project references.bib (creates if missing)
cat(pkg_bib, file = "references.bib", sep = "\n", append = TRUE)
cat("Package citations appended to references.bib\n")
```

**Important:** After merging, open `references.bib` and check for duplicate
entries. If you regenerate `packages.bib` later (e.g., after adding a new
package), merge only the new entries.

# Which Packages Should I Cite?

Per the [FORCE11 guidelines](https://doi.org/10.7717/peerj-cs.86) and
[rOpenSci advice](https://ropensci.org/blog/2021/11/16/how-to-cite-r-and-r-packages/):

| Always cite | Consider citing | Usually not necessary |
|-------------|-----------------|----------------------|
| Statistical method packages (`metafor`, `lme4`, `survival`) | Data-wrangling packages (`dplyr`, `tidyr`, `data.table`) | Low-level infrastructure (`rlang`, `cli`, `glue`) |
| Domain-specific packages (`tidyhydat`, `weathercan`) | Visualization packages (`ggplot2`, `plotly`) | Development tools (`devtools`, `testthat`) |
| Data-retrieval packages (`httr2`, `DBI`, `RSQLite`) | Reporting packages (`knitr`, `rmarkdown`, `gt`) | |

When in doubt, cite it. Package authors appreciate the recognition, and it
improves reproducibility.

# Further Reading

- LaZerte, S. (2021). [How to Cite R and R Packages](https://ropensci.org/blog/2021/11/16/how-to-cite-r-and-r-packages/). rOpenSci Blog.
- Smith, A. M., Katz, D. S., & Niemeyer, K. E. (2016). [Software Citation Principles](https://doi.org/10.7717/peerj-cs.86). PeerJ Computer Science, 2, e86.
- [grateful package](https://github.com/Pakillo/grateful) — Automated citation paragraph generation.
- [cffr package](https://docs.ropensci.org/cffr) — Create Citation File Format (CFF) files for R packages.
