---
title: "Summary Table Template (gtsummary)"
date: today
format:
  html:
    toc: true
    self-contained: true
---

<!--
  Template: summary-table.qmd
  Purpose:  Demonstrate publication-ready summary tables using gtsummary.
            gtsummary is the default package for descriptive/summary tables
            in the Research Workflow Assistant.

  Usage:
    1. Copy this file into your project directory.
    2. Replace the example data with your dataset.
    3. Customize the table specifications for your variables.
    4. Render with:  quarto render summary-table.qmd
-->

# Publication-Ready Summary Tables with gtsummary

The `gtsummary` package creates publication-ready analytical and summary tables
in R. It integrates with `gt` for rendering and supports export to Word, PDF,
HTML, and LaTeX. See the [gtsummary documentation](https://www.danieldsjoberg.com/gtsummary/) for full details.

## Setup

```{r}
#| label: setup
#| message: false

# Install if needed (uncomment):
# install.packages(c("gtsummary", "gt", "tidyverse"))

library(gtsummary)
library(gt)
library(tidyverse)
```

## Example 1: Demographic / Baseline Characteristics Table (Table 1)

The most common use — a "Table 1" summarizing participant characteristics by group.

```{r}
#| label: tbl-demographics
#| tbl-cap: "Baseline Characteristics of Study Participants"

# Using built-in trial dataset from gtsummary
trial |>
  select(trt, age, grade, stage, marker) |>
  tbl_summary(
    by = trt,                          # stratify by treatment group
    missing = "ifany",                 # show missing counts if any
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    label = list(
      age ~ "Age (years)",
      grade ~ "Tumor Grade",
      stage ~ "T Stage",
      marker ~ "Marker Level"
    )
  ) |>
  add_p() |>                           # add p-values
  add_overall() |>                     # add overall column
  modify_header(label ~ "**Characteristic**") |>
  bold_labels()
```

## Example 2: Regression Table

Summarize regression model results in publication format.

```{r}
#| label: tbl-regression
#| tbl-cap: "Logistic Regression: Predictors of Treatment Response"

glm(response ~ age + grade + stage, data = trial, family = binomial) |>
  tbl_regression(
    exponentiate = TRUE,              # show odds ratios
    label = list(
      age ~ "Age (years)",
      grade ~ "Tumor Grade",
      stage ~ "T Stage"
    )
  ) |>
  bold_p(t = 0.05) |>                # bold significant p-values
  add_global_p()                      # add global p for multi-level vars
```

## Example 3: Cross-Tabulation

```{r}
#| label: tbl-cross
#| tbl-cap: "Treatment Response by Tumor Grade"

trial |>
  tbl_cross(
    row = grade,
    col = response,
    percent = "row",
    label = list(grade ~ "Tumor Grade", response ~ "Response")
  ) |>
  add_p(source_note = TRUE)
```

## Example 4: Stacked / Merged Tables

Combine multiple tables for multi-outcome or multi-model presentations.

```{r}
#| label: tbl-stacked
#| tbl-cap: "Outcomes by Treatment Group"

# Table for continuous outcomes
t1 <- trial |>
  select(trt, marker, ttdeath) |>
  tbl_summary(
    by = trt,
    statistic = all_continuous() ~ "{median} ({p25}, {p75})",
    missing = "no"
  ) |>
  add_p()

# Table for categorical outcomes
t2 <- trial |>
  select(trt, response, death) |>
  tbl_summary(by = trt, missing = "no") |>
  add_p()

# Stack them
tbl_stack(list(t1, t2), group_header = c("Continuous Outcomes", "Categorical Outcomes"))
```

## Exporting Tables

```{r}
#| label: exports
#| eval: false

# To Word document (via flextable):
tbl <- trial |>
  select(trt, age, grade) |>
  tbl_summary(by = trt)

# Save as Word
tbl |> as_flex_table() |> flextable::save_as_docx(path = "table1.docx")

# Save as LaTeX (for PDF manuscripts)
tbl |> as_gt() |> gt::as_latex()

# Save as HTML
tbl |> as_gt() |> gt::gtsave("table1.html")
```

## Software Citation

```{r}
#| label: cite-packages
#| eval: false

# Generate BibTeX entries for gtsummary and gt
knitr::write_bib(c("gtsummary", "gt"), file = "packages.bib")

# Example Methods text:
# "Summary tables were created using the gtsummary package (v2.0.0;
#  Sjoberg et al., 2021) in R Statistical Software (v4.4.1; R Core
#  Team, 2025)."
```

## Alternative Packages

If `gtsummary` does not fit your needs:

- **`flextable`** — Best for Word (.docx) output; full control over cell formatting
- **`kableExtra`** — Lightweight; good for PDF/LaTeX with `knitr::kable()`
- **`huxtable`** — Flexible; supports conditional formatting, multi-format export
- **`gt`** — The rendering engine behind `gtsummary`; use directly for custom tables

For Python users, see `great_tables` (by Posit) or `itables` (interactive HTML).
