---
title: "Progress Brief"
subtitle: "[Project Title]"
author: "[PI Name]"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
format:
  docx:
    toc: false
  pdf:
    toc: false
    documentclass: article
    geometry: margin=1in
---

```{r}
#| include: false
# Load project tracking data
library(yaml)
project <- yaml::read_yaml("project-tracking/project.yaml")
tasks <- yaml::read_yaml("project-tracking/tasks.yaml")
decisions <- yaml::read_yaml("project-tracking/decisions.yaml")
```

# Project Status

**Current Phase:** `r project$current_phase`
**Start Date:** `r project$start_date`
**Target End:** `r project$target_end`

# Phase Progress

```{r}
#| echo: false
# Generate phase progress table
phases_df <- do.call(rbind, lapply(project$phases, function(p) {
  completed <- sum(sapply(p$milestones, function(m) m$status == "completed"))
  total <- length(p$milestones)
  data.frame(
    Phase = p$name,
    Status = p$status,
    Milestones = paste0(completed, "/", total),
    Target_End = p$target_end %||% "TBD"
  )
}))
knitr::kable(phases_df, col.names = c("Phase", "Status", "Milestones", "Target End"))
```

# Recent Accomplishments

```{r}
#| echo: false
completed_milestones <- unlist(lapply(project$phases, function(p) {
  sapply(p$milestones[sapply(p$milestones, function(m) m$status == "completed")],
         function(m) paste0("- ", m$name, " (", p$name, ")"))
}))
```

`r if(length(completed_milestones) > 0) paste(tail(completed_milestones, 5), collapse = "\n") else "No milestones completed yet."`

# Blockers

```{r}
#| echo: false
blocked_tasks <- tasks$tasks[sapply(tasks$tasks, function(t) t$status == "blocked")]
if(length(blocked_tasks) > 0) {
  cat(paste(sapply(blocked_tasks, function(t) 
    paste0("- **", t$description, "**: ", t$notes %||% "No details")), 
    collapse = "\n"))
} else {
  cat("No current blockers.")
}
```

# Recent Decisions

```{r}
#| echo: false
recent <- tail(decisions$decisions, 3)
if(length(recent) > 0) {
  cat(paste(sapply(recent, function(d) 
    paste0("- **", d$decision, "** (", d$date, "): ", d$rationale)), 
    collapse = "\n"))
} else {
  cat("No decisions recorded yet.")
}
```

# Next Steps

[To be filled in by the research team]

---

*This brief was generated with AI assistance. All project data was entered and verified by the research team.*
