---
title: "[Manuscript Title]"
author:
  - name: "[Lead Author, Credentials]"
    affiliation:
      - name: "[Institution / Organization]"
        city: "[City]"
        region: "[State / Province]"
        country: "[Country]"
        url: "[Organization URL]"
    email: "[author@example.org]"
    orcid: "[0000-0000-0000-0000]"
    url: "[Profile URL]"
    corresponding: true
  - name: "[Co-Author, Credentials]"
    affiliation:
      - name: "[Institution / Organization]"
        city: "[City]"
        region: "[State / Province]"
        country: "[Country]"
        url: "[Organization URL]"
    email: "[coauthor@example.org]"
    orcid: "[0000-0000-0000-0000]"
    url: "[Profile URL]"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
abstract: |
  [Structured or unstructured abstract, typically 150-300 words]
format:
  docx:
    toc: true
    toc-depth: 3
    number-sections: true
bibliography: references.bib
csl: apa.csl  # Default — agents replace with the project's configured style
execute:
  echo: false
  warning: false
---

<!-- If project-config.yaml defines research_assistant.authors, replace the placeholders in this front matter with those author records before rendering. -->
<!-- CSL style resolution: project output_defaults.csl → user default_citation_style → apa.csl fallback. Agents substitute the correct style when instantiating this template. -->
<!-- If RWA is cited in this document, add the `vanzyl2026rwa` entry from templates/rwa-citation.bib to references.bib. -->

# Introduction

[Background, rationale, and objectives. Move from broad context to specific gap to research question.]

# Methods

## Study Design

[Describe the study design and setting]

## Participants / Data Sources

[Describe population, eligibility criteria, data sources]

## Variables / Measurements

[Describe outcome variables, exposures, covariates]

## Statistical Analysis

[Describe analytical approach, software, handling of missing data]

<!-- Tip: cite the software you used for analysis. For R, use citation("pkg")
     or knitr::write_bib() to generate BibTeX entries. For Python, use the
     citation helper in analysis-templates/python/cite-python-packages.qmd.
     See analysis-templates/R/cite-r-packages.qmd for a full walkthrough.

     Example text:
     All analyses were performed using R Statistical Software (v4.4.1;
     R Core Team, 2025). Meta-analysis was conducted with the metafor
     package (v4.6-0; Viechtbauer, 2010). Data visualization used ggplot2
     (v3.5.1; Wickham, 2016). -->

## Use of Research Workflow Assistant (RWA)

[Describe whether Research Workflow Assistant (RWA), accessed through GitHub Copilot in VS Code, was used for literature searching, workflow organization, drafting support, analysis scripting, or citation management. Specify which tasks were AI-assisted, which decisions remained with the human authors, and how outputs were reviewed or verified before use. When RWA is described here, cite it as [@vanzyl2026rwa]. If you cite reporting standards, tools, or database attributions here, verify those references in references.bib before submission.]

## Ethical Considerations

[IRB/ethics approval, informed consent]

# Acknowledgments And AI Disclosure

[Acknowledge non-author contributors, funding, or institutional support as appropriate.]

**RWA disclosure:** Research Workflow Assistant (RWA), accessed through GitHub Copilot in VS Code, may have been used to assist with [describe specific tasks]. All study design, source selection, analytic decisions, interpretation, and final approval remained with the human authors. The authors reviewed, revised, and verified all AI-assisted content and take full responsibility for the final work.

If you mention RWA in the acknowledgments, cite it as [@vanzyl2026rwa].

# References

# Results

## Descriptive Statistics

```{r}
#| label: tbl-descriptive
#| tbl-cap: "Participant characteristics"
# Descriptive statistics table
```

## Primary Analysis

[Present main results]

```{r}
#| label: fig-main
#| fig-cap: "[Description of main figure]"
# Main results figure
```

## Secondary Analyses

[Additional analyses, subgroups, sensitivity]

# Discussion

## Summary of Findings

[Key findings in context of existing literature]

## Strengths and Limitations

[Methodological strengths and limitations]

## Implications

[Clinical/policy/research implications]

# Conclusions

[Brief summary]

# Acknowledgments

[Non-author contributors, funding]

**AI Tool Disclosure:** [Include if AI tools were used, per ICMJE Section II.A.4. See `compliance/ai-disclosure-template.md` for templates.]

**Data Source Attribution:** [If Semantic Scholar was used, include: "Literature search results were supplemented using the Semantic Scholar Open Data Platform (Kinney et al., 2023), which provides data under a CC BY-NC license." Add similar attributions for other databases as required by their terms of service.]

# References
