---
title: "[Manuscript Title]"
author:
  - name: "[First Author]"
    affiliation: "[Institution]"
    orcid: "[ORCID]"
    corresponding: true
  - name: "[Co-Author]"
    affiliation: "[Institution]"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
abstract: |
  [Structured or unstructured abstract, typically 150-300 words]
format:
  docx:
    toc: true
    toc-depth: 3
    number-sections: true
bibliography: references.bib
csl: apa.csl
execute:
  echo: false
  warning: false
---

# Introduction

[Background, rationale, and objectives. Move from broad context to specific gap to research question.]

# Methods

## Study Design

[Describe the study design and setting]

## Participants / Data Sources

[Describe population, eligibility criteria, data sources]

## Variables / Measurements

[Describe outcome variables, exposures, covariates]

## Statistical Analysis

[Describe analytical approach, software, handling of missing data]

## Ethical Considerations

[IRB/ethics approval, informed consent]

# Results

## Descriptive Statistics

```{r}
#| label: tbl-descriptive
#| tbl-cap: "Participant characteristics"
# Descriptive statistics table
```

## Primary Analysis

[Present main results]

```{r}
#| label: fig-main
#| fig-cap: "[Description of main figure]"
# Main results figure
```

## Secondary Analyses

[Additional analyses, subgroups, sensitivity]

# Discussion

## Summary of Findings

[Key findings in context of existing literature]

## Strengths and Limitations

[Methodological strengths and limitations]

## Implications

[Clinical/policy/research implications]

# Conclusions

[Brief summary]

# Acknowledgments

[Non-author contributors, funding]

**AI Tool Disclosure:** [Include if AI tools were used, per ICMJE Section II.A.4. See `compliance/ai-disclosure-template.md` for templates.]

**Data Source Attribution:** [If Semantic Scholar was used, include: "Literature search results were supplemented using the Semantic Scholar Open Data Platform (Kinney et al., 2023), which provides data under a CC BY-NC license." Add similar attributions for other databases as required by their terms of service.]

# References
