---
title: "[Title of Systematic Review]"
subtitle: "A Systematic Review [and Meta-Analysis]"
author:
  - name: "[First Author]"
    affiliation: "[Institution]"
    orcid: "[ORCID]"
    corresponding: true
  - name: "[Second Author]"
    affiliation: "[Institution]"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
abstract: |
  **Background:** [Context and rationale]
  **Objectives:** [Review question and objectives]
  **Methods:** [Search strategy, databases, eligibility criteria, data extraction, synthesis methods]
  **Results:** [Number of studies, key findings, effect estimates]
  **Conclusions:** [Interpretation, implications]
  **Registration:** [PROSPERO/OSF number]
format:
  docx:
    toc: true
    toc-depth: 3
    number-sections: true
bibliography: references.bib
csl: apa.csl
execute:
  echo: false
  warning: false
---

# Introduction

## Rationale

[Describe the rationale for the review in the context of what is already known]

## Objectives

[Provide an explicit statement of the review question(s) using PICO/PEO framework]

# Methods

## Protocol and Registration

This systematic review was conducted in accordance with the PRISMA 2020 guidelines [@page2021prisma]. The protocol was registered with [PROSPERO/OSF] (registration number: [number]).

## Eligibility Criteria

### Inclusion Criteria

[Describe using PICO/PEO elements]

### Exclusion Criteria

[List exclusion criteria]

## Information Sources

[Describe all databases searched, with dates]

## Search Strategy

[Present complete search strategy for at least one database. Full strategies for all databases are available in Supplementary Material.]

## Selection Process

[Describe screening process: number of reviewers, independence, tools, disagreement resolution]

## Data Collection Process

[Describe data extraction methods]

## Data Items

[List variables extracted]

## Study Risk of Bias Assessment

[Describe RoB tool and process]

## Effect Measures

[Describe effect measures used]

## Synthesis Methods

[Describe qualitative and quantitative synthesis methods]

## AI Tool Disclosure

[Tool name] was used to assist with [specific tasks]. All screening decisions, data extraction, and risk of bias assessments were performed independently by the research team. See AI Contributions Log (Supplementary Material) for complete details.

# Results

## Study Selection

[Describe selection process with PRISMA flow diagram]

```{r}
#| label: fig-prisma
#| fig-cap: "PRISMA 2020 flow diagram"
# PRISMA flow diagram will be generated from prisma-tracker data
# Use the prisma-tracker MCP server's generate_prisma_diagram() output
```

## Study Characteristics

[Describe included studies]

```{r}
#| label: tbl-characteristics
#| tbl-cap: "Characteristics of included studies"
# Table of study characteristics
```

## Risk of Bias in Studies

[Present RoB assessments]

## Results of Individual Studies

[Present individual study results]

## Results of Syntheses

[Present synthesis results, forest plots if meta-analysis performed]

```{r}
#| label: fig-forest
#| fig-cap: "Forest plot of [outcome]"
# Forest plot code (using metafor or meta package)
```

## Reporting Biases

[Assess reporting biases, funnel plot if applicable]

## Certainty of Evidence

[Present GRADE or equivalent assessment]

# Discussion

## Summary of Evidence

[General interpretation in context of other evidence]

## Limitations

[Limitations of evidence and review process]

## Implications

[Implications for practice, policy, and research]

# Conclusions

[Brief summary of findings and their implications]

# Acknowledgments

[Acknowledge non-author contributors]

**AI Tool Disclosure:** [Tool name and version] was used during the preparation of this manuscript to [specific uses]. The authors reviewed and edited all AI-assisted content and take full responsibility for the content of the publication.

# References
